---
title: "BIOSTAT625 project test"
author: "Hengde Ouyang"
date: "2022-11-27"
output: 
  pdf_document:
    latex_engine: xelatex
---

# Read the Data and Remove the row contains NA
```{r}
raw_data = read.csv("raw1.csv")
# Remove the row contains NA
complete_data = na.omit(raw_data)
```

# Transform some of the data
```{r,warning=FALSE}
transform_data = complete_data
# If the person is female, we code 0
transform_data$Sex = ifelse(transform_data$Sex=="Female",0,1)

# Survival months to numeric, and remove the unknown
# This is our response variable
transform_data$Survival.months = as.numeric(transform_data$Survival.months)
transform_data = na.omit(transform_data)

# Remove "years" in Single Age
# Code, if you want to remove:
# pattern = "years"
# transform_data$Age.recode.with.single.ages.and.85.  #=gsub(transform_data$Age.recode.with.single.ages.and.85.,pattern=pattern,
# replacement = "")

# Tumor Size to numeric, and remove the Blank
transform_data$CS.tumor.size..2004.2015. = as.numeric(transform_data$CS.tumor.size..2004.2015.)
transform_data = na.omit(transform_data)

# Tumor number to numeric, and remove the Blank
transform_data$Total.number.of.in.situ.malignant.tumors.for.patient = as.numeric(transform_data$Total.number.of.in.situ.malignant.tumors.for.patient)
transform_data = na.omit(transform_data)

### Silly Code
transform_data$Age.recode.with..1.year.olds = factor(transform_data$Age.recode.with..1.year.olds)
transform_data$Primary.Site...labeled = factor(transform_data$Primary.Site...labeled)
transform_data$Derived.AJCC.Stage.Group..6th.ed..2004.2015. = factor(transform_data$Derived.AJCC.Stage.Group..6th.ed..2004.2015.)
transform_data$ER.Status.Recode.Breast.Cancer..1990.. = factor(transform_data$ER.Status.Recode.Breast.Cancer..1990..)
transform_data$PR.Status.Recode.Breast.Cancer..1990.. = factor(transform_data$PR.Status.Recode.Breast.Cancer..1990..)
transform_data$Survival.months.flag = factor(transform_data$Survival.months.flag)
transform_data$Race.ethnicity = factor(transform_data$Race.ethnicity)


```


# Summary Statistics
```{r}
summary(transform_data)
```


# Principal Components Regression
Suppose we have the original predictors $X_{1},X_{2},...,X_{p}$,let 
$Z_{1},Z_{2},...,Z_{p}$ represent $M<p$ linear combinations of our original
p predictors, that is:
$$Z_{m}=\sum_{j = 1}^{p} \phi_{jm}X_{j}$$


We show an example of Princinpal Components Regression:
```{r,warning=FALSE}
#Before running the code, you need the "pls" and "ISLR" package
#install.packages("pls")
library(pls)
library(ISLR)

```


```{r}
set.seed(2)
Hitters =na.omit(Hitters)
pcr.fit=pcr(Salary~.,data=Hitters, scale=TRUE,
validation ="CV")
summary (pcr.fit)
```

```{r}
validationplot(pcr.fit,val.type="MSEP")
```

training and testing data:
```{r}
set.seed(1)
train=sample(c(TRUE ,FALSE), nrow(Hitters ),rep=TRUE)
test=(!train)


x=model.matrix(Salary~.,Hitters)[,-1]
y=Hitters$Salary

y.test=y[test]
```


```{r}
set.seed(2)
pcr.fit=pcr(Salary~.,data=Hitters,subset=train,scale=TRUE,
validation ="CV")
pcr.pred=predict (pcr.fit,x[test ,],ncomp =7)
mean((pcr.pred -y.test)^2)
pcr.pred=predict (pcr.fit,x[test ,],ncomp =2)
mean((pcr.pred -y.test)^2)
pcr.pred=predict (pcr.fit,x[test ,],ncomp =1)
mean((pcr.pred -y.test)^2)
```




# Random Forest




```{r,warning=FALSE}
#Before running the code, you need the "randomForest" package
#install.packages("randomForest")
library(randomForest)

```


```{r}
set.seed(3)
bag.Hitters =  randomForest(Salary~.,data=Hitters, subset=train,
mtry=13,importance =TRUE)

yhat.bag = predict(bag.Hitters,newdata=Hitters[test,])
mean((yhat.bag -y.test)^2)
```

# Comparision of time
```{r}
system.time({
pcr.fit=pcr(Salary~.,data=Hitters,subset=train,scale=TRUE,
validation ="CV")
})
```

```{r}
system.time({
bag.Hitters =  randomForest(Salary~.,data=Hitters, subset=train,
mtry=13,importance =TRUE)
})
```


# Try to use future package
```{r}
library(future)
plan(multisession)
set.seed(3)
system.time({
bag.Hitters2 =  future(randomForest(Salary~.,data=Hitters, subset=train,
mtry=13,importance =TRUE),seed = TRUE)
})
```

```{r}
yhat.bag2 = predict(value(bag.Hitters2),newdata=Hitters[test,])
mean((yhat.bag2 -y.test)^2)
```






